---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tjspack

<!-- badges: start -->
<!-- badges: end -->

The goal of tjspack is to assist in data cleaning for longitudinal study datasets. Functions within are intended to ensure
demographic data doesn't change for subjects on different visits, dates visit numbers, and ages are entered correctly. As well as
checking if things are correctly ordered chronologically

## Installation

You can install the development version of tjspack like so:

``` r
library(devtools)
install_github("tylerserm/tjspack")
```

## Example

This is a basic example which details how one can use the check_age function.
Note that for ID 1, the age is entered incorrectly on the third visit

```{r example}
library(tjspack)

example1 <- data.frame(
ID = c(1, 1, 1, 2, 2),
TestDate = as.Date(c("2023-01-01", "2023-01-01", "2023-04-01", "2023-02-01", "2023-02-01")),
DOB = as.Date(c("2017-06-01", "2017-06-01", "2017-06-01", "2017-12-14", "2017-12-14")),
Age = c(5.58, 5.58, 8.83, 5.05, 5.05))

check_age(example, "ID", "TestDate", "DOB", "Age")

example1
```
